Behavior Driven Development in Haskell

### 2012-05-28 paper.tex
evidences -> evidence.

In BDD, all behavior should be documented, together with some
\emph{evidence}\/s for that behavior.
For this purpose, 
we brought a test method from other programming language community and
extended missing features.
In particular, we introduced Doctest~\cite{doctest} and
integrated it with QuickCheck~\cite{quickcheck}
for evidences for users.
To specify evidences for developers,
we make use of Hspec inspired by RSpec~\cite{rspec}
and we proposed some combinators to it.

@@

We can categorize evidences into two types: for users and
and for developers.
Evidences for users are those which are useful as user documentation.
By contrast, evidences for developers are
those which are not useful as user documentation
but should be written as code documentation.
A good example of evidences for developers
is a corner example registered in a bug tracking system.

BDD tools should be treat both of them and provide functionality
to extract evidences for users as user documentation.
If we can archive this, BDD tools can be used for
triple purpose: design, documentation and automatic testing.

@@

A good candidate to describe evidences for users is Doctest in Python community.
It extends Python's documentation strings so that they
can contain examples.
Examples can be tested by the Python interpreter.

@@

To describe evidences for developers, we can use Hspec.
Hspec provides an EDSL for defining specs.  A spec organizes evidence in a
tree structure, and it is defined with \texttt{it} and \texttt{describe}.

@@

A spec can be run with the \texttt{hspec} function.  When a spec is run, all
evidences are evaluated, and failing evidences are reported to developers.

@@

\begin{itemize}
\item
    The Haskell community does not have common syntax to describe
    evidences with Haddock markup syntax.
\item
    The Haskell community does not have a BDD tool for
    evidences for users.
\item
    Hspec does not provide combinators to express expectations with
    BDD terminology
\end{itemize}

@@

To be able to specify evidences for users, we propose to add new
markup to Haddock and to use Doctest integrated with QuickCheck. To
describe evidences for developers, we propose to use Hspec also
integrated with QuickCheck.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Evidences for users}
\label{sec:for-users}

Evidences for users should be documented with Haddock markup syntax.
For this purpose, we propose the followings:

@@

\noindent Evidences in comments are automatically
tested by the Doctest command. We also provided a library
to do this job. Its implementation will discussed in Section~\ref{sec:doctest}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Evidences for developers}

We propose following extensions for Hspec to 
describe evidences for developers.

@@

As we described, Haddock version 2.7.2 or earlier provides markup syntax for
code blocks, but does not provide any markup syntax for evidences.
We implemented markup support for examples ({\tt >>>}) and it is available in
Haddock version 2.8 or later.
We also extended Haddock to support markup for properties ({\tt prop>}) which
will be available with a future release of Haddock.
Both examples and properties in Haddock comments are pretty-printed when
generating documentation, (e.g. as HTML).

@@

TBD(@sol): how to obtain evidences from Haddock markup.
Old version uses Haddock API but now uses GHC API directly. Why?

@@

\noindent Note that the {\tt placeholders} library provide better keyword such as {\tt notImplemented} and {\tt todo}.
Next we write evidences for users in Haddock markup:

@@

\noindent We should execute Doctest to ensure no syntax errors.
Also, we specifies evidences for developers with Hspec:

>>>>

In BDD, all behavior should be documented, together with some
\emph{evidence}\/s for that behavior.
For this purpose, 
we brought a test method from other programming language community and
extended missing features.
In particular, we introduced Doctest~\cite{doctest} and
integrated it with QuickCheck~\cite{quickcheck}
for evidence for users.
To specify evidence for developers,
we make use of Hspec inspired by RSpec~\cite{rspec}
and we proposed some combinators to it.

@@

We can categorize evidence into two types: for users and
and for developers.
Evidence for users are those which are useful as user documentation.
By contrast, evidence for developers are
those which are not useful as user documentation
but should be written as code documentation.
A good example of evidence for developers
is a corner example registered in a bug tracking system.

BDD tools should be treat both of them and provide functionality
to extract evidence for users as user documentation.
If we can archive this, BDD tools can be used for
triple purpose: design, documentation and automatic testing.

@@

A good candidate to describe evidence for users is Doctest in Python community.
It extends Python's documentation strings so that they
can contain examples.
Examples can be tested by the Python interpreter.

@@

To describe evidence for developers, we can use Hspec.
Hspec provides an EDSL for defining specs.  A spec organizes evidence in a
tree structure, and it is defined with \texttt{it} and \texttt{describe}.

@@

A spec can be run with the \texttt{hspec} function.  When a spec is run, all
evidence are evaluated, and failing evidence are reported to developers.

@@

\begin{itemize}
\item
    The Haskell community does not have common syntax to describe
    evidence with Haddock markup syntax.
\item
    The Haskell community does not have a BDD tool for
    evidence for users.
\item
    Hspec does not provide combinators to express expectations with
    BDD terminology
\end{itemize}

@@

To be able to specify evidence for users, we propose to add new
markup to Haddock and to use Doctest integrated with QuickCheck. To
describe evidence for developers, we propose to use Hspec also
integrated with QuickCheck.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Evidence for users}
\label{sec:for-users}

Evidence for users should be documented with Haddock markup syntax.
For this purpose, we propose the followings:

@@

\noindent Evidence in comments are automatically
tested by the Doctest command. We also provided a library
to do this job. Its implementation will discussed in Section~\ref{sec:doctest}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Evidence for developers}

We propose following extensions for Hspec to 
describe evidence for developers.

@@

As we described, Haddock version 2.7.2 or earlier provides markup syntax for
code blocks, but does not provide any markup syntax for evidence.
We implemented markup support for examples ({\tt >>>}) and it is available in
Haddock version 2.8 or later.
We also extended Haddock to support markup for properties ({\tt prop>}) which
will be available with a future release of Haddock.
Both examples and properties in Haddock comments are pretty-printed when
generating documentation, (e.g. as HTML).

@@

TBD(@sol): how to obtain evidence from Haddock markup.
Old version uses Haddock API but now uses GHC API directly. Why?

@@

\noindent Note that the {\tt placeholders} library provide better keyword such as {\tt notImplemented} and {\tt todo}.
Next we write evidence for users in Haddock markup:

@@

\noindent We should execute Doctest to ensure no syntax errors.
Also, we specifies evidence for developers with Hspec:
